Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-26 doc/RMD/mlg-mcg.Rmd
update MCG analysis to only include truely unconnected MLGs

Of interest is accounting for the communities that have a single MLG within the
community. This gives us a lower bound at how many MLGs are reproductively
isolated. We detect these by filtering for communities that only have one MLG:

```{r}
singles <- the_communities %>% 
  mutate(MLG = grepl("MLG", vertex)) %>% 
  group_by(community) %>% 
  summarize(n = sum(MLG)) %>% 
  filter(n == 1) %>%
  inner_join(the_communities, by = "community") %>% 
  mutate(type = case_when(grepl("MLG", vertex) ~ "MLG", TRUE ~ "MCG")) %>% 
  group_by(community, type) %>% 
  summarize(vertex = paste(as.character(vertex), collapse = ", ")) %>%
  spread(type, vertex) 
# assessing that this truely contains the singletons
nrow(inner_join(singles, the_communities, by = c("MLG" = "vertex")))
knitr::kable(singles)
```

@@

make_rgb <- function(x, alpha = 1){
  out <- col2rgb(x, alpha = TRUE)
  out[4, ] <- floor(out[4, ]*alpha)
  out      <- apply(out, 2, paste, collapse = ", ")
  paste0("rgba(", out, ")")
}
vg <- g %>%
  set_vertex_attr("size", value = osize) %>%
  set_vertex_attr("value", value = osize) %>%
  set_vertex_attr("color", value = viridis(4)[as.integer(factor(the_communities$comm))]) %>%
  set_vertex_attr("shape", value = ifelse(V(.)$type == "Multilocus Genotype", "triangle", "circle")) %>%
  set_edge_attr("width", value = E(.)$weight) %>%
  toVisNetworkData()
vg$nodes <- vg$nodes %>% 
  group_by(id) %>%
  mutate(color = list(list(background = make_rgb(color, 0.8), 
                      border = make_rgb(rep("black", n()), 0.8),
                      highlight = list(background = make_rgb(color),
                                       border = make_rgb(rep("black", n()))
                                       )
                      )))
vgn <- visNetwork(nodes = vg$nodes, edges = vg$edges,# height = "500px", 
                  main = "Relation of Multilocus Genotypes and MCGs")
set.seed(2017-05-03)
vgn
# vgn %>%
#   visIgraphLayout("layout_nicely") %>%
#   visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 
#              nodesIdSelection = TRUE)
#   
```

>>>>

Of interest is accounting for the communities that have a single MLG within the
community. This gives us a lower bound at how many MLGs are reproductively
isolated. We detect these by filtering for communities that only have one MLG.
But, because we do have communities with one MLG that are peripheral to larger
communities, we can also check the degree of the MCG and MLG nodes to ensure 
that they correspond one-to-one.

```{r}
singles <- the_communities %>% 
  mutate(MLG = grepl("MLG", vertex)) %>% 
  group_by(community) %>% 
  summarize(n = sum(MLG)) %>% 
  filter(n == 1) %>%
  inner_join(the_communities, by = "community") %>% 
  mutate(type = case_when(grepl("MLG", vertex) ~ "MLG", TRUE ~ "MCG")) %>% 
  group_by(community, type) %>% 
  summarize(vertex = paste(as.character(vertex), collapse = ", ")) %>%
  spread(type, vertex) %>%
  mutate(MCG_degree = map_dbl(strsplit(MCG, ", "), . %>% degree(g, .) %>% sum())) %>%
  mutate(MLG_degree = degree(g, MLG)) %>%
  filter(MLG_degree - MCG_degree == 0)
# assessing that this truely contains the singletons
nrow(inner_join(singles, the_communities, by = c("MLG" = "vertex")))
singles

knitr::kable(singles)
```

This result indicates that there are `r nrow(singles)` MLGs that are unable to
pair with any other MLGs due to mycelial incompatibility. This represents 
`r paste0(sprintf("%0.2f", nrow(singles)/nmll(dat11, "original")*100), "%")` 
of the total MLGs.

```{r}
mll(dat11) <- "original"
mlg.table(dat11, ~Region, color = TRUE, mlgsub = as.integer(gsub("MLG.", "", singles$MLG)))
p <- last_plot()
p + 
  scale_fill_brewer(palette = "Set1") +
  ggtitle("Incompatible Multilocus Genotypes") +
  labs(list(caption = "These MLGs cannot pair with any other MLG due to\nmycelial incompatibility"))
```

@@

make_rgb <- function(x, alpha = 1){
  out <- col2rgb(x, alpha = TRUE)
  out[4, ] <- floor(out[4, ]*alpha)
  out      <- apply(out, 2, paste, collapse = ", ")
  paste0("rgba(", out, ")")
}
vg <- g %>%
  set_vertex_attr("size", value = osize) %>%
  set_vertex_attr("value", value = osize) %>%
  set_vertex_attr("color", value = viridis(4)[as.integer(factor(the_communities$comm))]) %>%
  set_vertex_attr("shape", value = ifelse(V(.)$type == "Multilocus Genotype", "triangle", "circle")) %>%
  set_edge_attr("width", value = E(.)$weight) %>%
  toVisNetworkData()
vg$nodes <- vg$nodes %>% 
  group_by(id) %>%
  mutate(color = list(list(background = make_rgb(color, 0.8), 
                      border = make_rgb(rep("black", n()), 0.8),
                      highlight = list(background = make_rgb(color),
                                       border = make_rgb(rep("black", n()))
                                       )
                      )))
vgn <- visNetwork(nodes = vg$nodes, edges = vg$edges,# height = "500px", 
                  main = "Relation of Multilocus Genotypes and MCGs")
set.seed(2017-05-03)

vgn %>%
  visIgraphLayout("layout_nicely") %>%
  visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 
             nodesIdSelection = TRUE)

### 2017-07-26 doc/manuscript/manuscript.Rmd
update MCG analysis to only include truely unconnected MLGs

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 26 of which consisted of a single MLG indicating that 15.76% of the 165 MLGs are unable to cross with any other MLG in this data set.
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
Overall, there was no clear pattern to the association between MLG and MCGs.

>>>>

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
Overall, there was no clear pattern to the association between MLG and MCGs.
